Understanding the Nanoscale Structure of Inverted Hexagonal Phase Lyotropic Liquid Crystal Polymer Membranes

### 2018-10-08 Ben_Manuscripts/structure_paper/Final_Draft.tex
added missing files

\title{Understanding the Nanoscale Structure of Inverted Hexagonal Phase Lyotropic Liquid Crystal Polymer Membranes}
\author{Benjamin J. Coscia}
\affiliation{Department of Chemical and Biological Engineering, University of Colorado Boulder, Boulder, CO 80309, USA}
\author{Joseph Yelk}
\author{Matthew A. Glaser}
\affiliation{Department of Physics, University of Colorado Boulder, Boulder CO, 80309, USA}
%\author{Richard D. Noble}
\author{Douglas L. Gin}
\affiliation{Department of Chemical and Biological Engineering, University of Colorado Boulder, Boulder, CO 80309, USA}
\alsoaffiliation{Department of Chemistry and Biochemistry, University of Colorado Boulder, Boulder CO 80309, USA}
\author{Xunda Feng}
\affiliation{Department of Chemical and Environmental Engineering, Yale University, New Haven, Connecticut 06511, USA}
\author{Michael R. Shirts}
\email{michael.shirts@colorado.edu}
\affiliation{Department of Chemical and Biological Engineering, University of Colorado Boulder, Boulder, CO 80309, USA}
\begin{document}
%MRS13: we might end up adjusting the author list a bit; I will ask Matt where he thinks he and Joe should go.  I will talk to Rich when we get this going; not clear that he actually has contributed.  But that's my job. 
  \graphicspath{{./figures/}}

@@

@@

  %MRSF: I changed the description of the ``demonstrating'', ``showing'' verbs to present tense.  The experiments are past tense, but the ``showing'' is present tense, since it is occuring while the reader is reading (like in a presentation - you are showing as they listen).
  \end{abstract}

@@

  % BJC5: Note to self -- how to use latex xr : Figure S~\ref{S-fig:monomer_color_coded}

  More highly selective nanoporous membranes would be extremely useful 
  %in order to perform 
  in performing complex aqueous separations with seawater and various 
  %sources 
  types of
  wastewater. 
  For exaple, 
  Sodium chloride and boron in seawater
  \cite{fritzmann_state---art_2007} and organic micropollutants found in
  municipal and industrial wastewaters \cite{schwarzenbach_challenge_2006} represent just a few of the diverse contaminants 
  of water sources. By efficiently
  separating contaminants from feed solutions with highly selective membranes,
  it is possible to reduce the number of required membrane passes and post-treatment steps
  needed for a given filtration process \cite{werber_materials_2016}, thus
  lowering energy requirements. Additionally, one can also extract valuable resources
  from the feed streams. For example, flowback water produced during hydraulic
  fracturing of shale formations contains dissolved species such as acetate whose
  extraction has economic value \cite{dischinger_application_2017}.

%MRSF: a few selelected edits above, but I think strikes the right balance.
% More highly selective membranes would be extremely useful for both the
% recovery of valuable dissolved species from and the purification of complex
%  aqueous solutions. The supply of fresh water is being increasingly strained
%  by population growth, industrialization and climate change
%  \cite{elimelech_future_2011}. Consequently, people are turning towards
%  treatment of unconventional water sources such as seawater and wastewater. 
%  In addition to obtaining fresh water from them, these water sources, 
%  particularly municipal and industrial wastewater, have been found to contain 
%  valuable resources which can be separated using highly selective membranes~\cite{sales_resource_2015}.
  % make possible with selectively permeable membranes -- better than thermal distillation

 % BJC14: (2) existing nanoporous membranes + drawbacks

  Reverse osmosis (RO) and nanofiltration (NF) are two prevailing membrane
  filtration processes that can be used to separate solutes on the order of 1 nm
  in size and smaller, including 
  %as well as 
  %MRSF: this does seem to blend RO and NF together; is this correct, or are there important differences that should be outlined?
  ions. However, the porous architecture of
  commercially available RO and NF membranes lack uniformity, which severely impacts their
  selectivity \cite{van_der_bruggen_review_2003}. Although scalable, their
  fabrication involves the spontaneous assembly of polymers into disordered
  structures that offer separation pathways that are tortuous and polydisperse
  in size \cite{werber_materials_2016}. Tortuosity increases the effective
  length that a solute must travel, while pore size polydispersity limits membrane
  selectivity.

  % BJC14: (3) justification for why nanostructured membranes are a good idea for this
  % long-term route towards highly selective membranes are those that are nanostructured

@@

%BJC14: zeolites aren't really nanostructured
% Zeolites have sub-nm
%  pores with a narrow pore size distribution, and MD simulations of these
%  materials show that they exhibit complete rejection of solvated ions
%  \cite{murad_molecular_1998}- however, experimental rejection was low and
%  attributed to interstitial defects formed during membrane synthesis
%  \cite{li_desalination_2004}.

%  Reverse osmosis (RO) and nanofiltration (NF) are two prevailing membrane
%  filtration processes that can be used to separate solutes on the order of 1 nm
%  in size and smaller, as well as ions. Both apply hydraulic pressure to the feed
%  solution in order to overcome osmotic pressure and force water and unfiltered
%  components through the membrane. RO membranes typically have a thin film
%  composite architecture which, at minimum, consists of a porous support layer
%  and a dense polymer active layer. Separation occurs based on the differences in
%  solubility and diffusivity of feed solution constituents in the active layer.
%  Nanofiltration membranes have explicit pores on the order of 1 nm in diameter
%  which are typically tortuous and polydisperse in size. They separate primarily
%  based on size exclusion as well as Donnan exclusion if the membrane has a
%  surface charge~\cite{van_der_bruggen_review_2003}.

  %MRS14: is selectivity towards water important, or selectivity against what one wants filtered out?
%  The energy consumed during the purification of seawater and wastewater 
%  using RO and NF membranes can be reduced by increasing their selectivity 
%  towards water \cite{werber_materials_2016}. While researchers often focus on
%  improving membrane permeability, the flux of water through a given membrane,
%  it has been shown that increased water permeability only slightly decreases
%  energy requirements \cite{cohen-tanugi_quantifying_2014}.
%  since the main energy requirement for a membrane separation is that needed
%  to create the hydraulic pressure necessary in order to overcome the 
%  osmotic pressure of the feed solution\cite{cohen-tanugi_quantifying_2014}. 
%  Often times, multiple membrane passes are required in order to sufficiently
  %MRS14: not sure the boric acid part is that relevant, since we don't really talk about 
  %that in the rest of the paper.
%  reduce contaminant levels. For example, boric acid is toxic to humans and 
%  is present in seawater at levels well above acceptable drinking standards.
%  It is difficult to sufficiently separate boric acid in a single pass, meaning
%  multiple passes are needed and energy requirements nearly double \cite{fritzmann_state---art_2007}.
%  This is necessary in the case where high-purity water
%  is needed \cite{singh_production_2009}, when separating small neutral solutes \cite{ozaki_rejection_2002} or
%  when removing boric acid from seawater \cite{fritzmann_state---art_2007}.  
%  The most effective way to reduce energy consumption
%  would be to reduce the number of membrane passes for a given system using 
%  more selective membranes. 
%   More highly selective membranes have the potential
%  to reduce the number of membrane passes in addition to eliminating the need 
%  for other post-treatment steps necessitated by incomplete separations.
  %dissolved species which negatively
%  affect human health.

%  Not only will increased selectivity reduce the energy consumption of membrane
%  processes, but it offers a way to recover valuable resources from the feed 
%  streams. Municipal and industrial wastewaters are contaminated by harmful
%  micropollutants, but also contain valuable dissolved solutes. Micropollutants include
%  pharmaceutical and personal care products, hormones, pesticides and industrial
%  chemicals, all of which have adverse effects on human health even at low 
%  concentrations\cite{schwarzenbach_challenge_2006}. A selective membrane has the % There is a need for membranes
%  potential to reject these components while still allowing the passage of valuable 
%  resources for further processing and recovery.

%  Municipal wastewaters are rich in carbon, nitrogen and phosphorus-containing 
%  compounds. The recovery of such products, which can be achieved using a 
%  selective membrane, has numerous potential uses \cite{sales_resource_2015}. 
%  For example, nitrogen and phosphorus recovery can help sustain fertilizer 
%  production which will help meet global food demand as population continues
%  to increase \cite{xie_membrane-based_2016}.

  %MRS14: could possibly be food details, if selective membranes are the key to purifying industrial waste waters. 
%  Industrial waste-waters are often quite complex with up to six times more 
%  total dissolved solids than seawater\cite{werber_materials_2016}. For example, flowback water, produced
%  during hydraulic fracturing of shale formations consists of relatively high
%  concentrations of salts, metals, and soluble organic compounds. The majority
%  of this water is disposed through deep well injection, however there is a growing public concern about
%  its management which has prompted the use of separation technologies such as 
%  RO and NF in order to reduce the volume of contaminated water~\cite{gregory_water_2011}.
%  Some of the dissolved organic compounds in flowback water, such as acetate, 
%  are potentially valuable and can be recovered with highly selective membranes
%  ~\cite{dischinger_application_2017}.

%  More highly selective membranes would be extremely useful for the recovery of
%  dissolved species in complex aqueous and organic solutions. For example,
%  selective membranes can further progress forward osmosis processes by reducing
%  the cost of separating potable water from draw
%  solutions~\cite{mccutcheon_novel_2005}. Additionally, flowback water (FW)
%  produced during hydraulic fracturing is a complex wastewater full of
%  potentially valuable dissolved organic compounds such as acetate that can be
%  recovered using highly selective membranes \cite{dischinger_application_2017}.
%  Finally, a significant amount of nitrogen and phosphorus from fertilizer is
%  found in wastewater. Selective membranes offer a route towards low energy
%  nutrient recovery from these wastewaters which can help sustain fertilizer
%  production and subsequently food production into the future
%  \cite{xie_membrane-based_2016}.
  % as we near peak production of minable phosphorus rock .

%  There is increasing pressure to reuse
%  hydraulic fracturing water rather than dispose of it in order to reduce social
%  and environmental impacts as well as cost \cite{theodori_hydraulic_2014}.
%  Rather fully than dispose of the wastestream generated in the recycling
%  process, we can instead use highly selective membranes in order to successfully
%  recover useful compounds.

%  The permeability-selectivity tradeoff is a well-known problem in the membrane
%  separations community. It is difficult to increase the permeability of a 
%  desired molecular or atomic species, while maintaining the same retention of
%  an undesired species.

%  Current commercial RO and NF membranes suffer limitations to their selectivity
%  because their structures incorporate some degree of randomness. Although scalable, their 
%  fabrication involves the spontaneous assembly of polymers into disordered structures
%  which offer separation pathways that are tortuous and polydisperse in size 
%  \cite{werber_materials_2016}. Tortuosity increases the effective length that a 
%  solute must travel, while polydispersity limits membrane selectivity.

%  This makes overcoming the well-known 
%  permeability-selectivity tradeoff a challenge. Namely, it is difficult to increase
%  the permeability of a desired molecular or atomic species, while maintaining the 
%  same retention of an undesired species \cite{werber_materials_2016}.

%MRS5: under solution-diffusion, then the free energy to partition into the membrane is important as well, correct?
%  Selective separation by a semipermeable membrane barrier can be broken into two
%  steps: barrier entry and transport through the barrier. A solute's ability to
  % BJC: can also come up with a third step: exit. Which I think is mostly chemical potential gradient
%  enter a given membrane barrier is decided by its size, shape, charge, polarity
%  and how those factors combine to interact with the membrane material. The same
%  properties affect the rate at which a solute travels across the barrier. Since
%  undesired solutes may still enter the membrane, their diffusion rates relative
%  to desired solutes are important
%  \cite{gin_polymerized_2008,wijmans_solution-diffusion_1995}. To achieve optimum
%  selectivity, it is necessary to design membranes in a way that one can tune the
%  relative diffusivities of desired and undesired solutes.

%BJC14: probably unnecessary?
%  Selective separation by a semipermeable membrane barrier is a function of the
%  geometric and chemical interactions of solutes with the membrane material. A
%  molecule's size, shape, charge and polarity combine to determine the degree to
%  which a solute partitions into a membrane and how fast it travels through the
%  membrane. To separate a component from a mixture, one must understand how to
%  design membranes in order to tune the relative transport rates of desired and
%  undesired solutes\cite{gin_polymerized_2008,wijmans_solution-diffusion_1995}.

% BJC: Maybe I don't even need to dive into explaining RO and NF.
%  Reverse osmosis (RO) and nanofiltration (NF) are membrane-based separation
%  processes widely used to create potable or reusable water by removing salts and
%  other dissolved compounds from water sources. RO membranes are typically thin
%  film composite (TFC) membranes with a porous support layer and a dense
%  crosslinked polymer active layer which  separates all solutes from water. NF
%  membranes have porous architectures with pores on the order of 1 nm in size
%  which separate primarily based on solute size
%  \cite{van_der_bruggen_review_2003}.

  % BJC14: (4) justify LLCs

@@

  The H\textsubscript{II}-phase pore geometry (Figure~\ref{fig:assembly}) has a
  higher theoretical capacity for transport than the Q\textsubscript{I} phase.
  The H\textsubscript{II} phase forms at room temperature in the presence of
  ca.~10 wt\% water and consists of hexagonally packed, hydrophilic pore
  columns\cite{smith_ordered_1997}. In the absence of water, neat monomer will
  form the same hexagonal columnar structure which, in the literature, has been
  referred to as the Col\textsubscript{h} thermotropic
  phase\cite{feng_scalable_2014}. 
  % BJC6: maybe not a necessary detail
%  The most
%  promising Q\textsubscript{I} phase for membrane applications forms at 70\degree
%  C when monomer is mixed with c.a.~20 wt\% glycerol\cite{carter_glycerol-based_2012}.
  Q\textsubscript{I}-phase membranes consist of a tortuous network of three
  dimensionally interconnected pores that prevent optimal through-plane
  transport. In contrast, the densely packed, non-tortuous and uniform sized
  pores of H\textsubscript{II}-phase membranes represent the ideal geometry for
  achieving high solute flux\cite{matyka_tortuosity-porosity_2008}.  However, the
  hexagonally packed LC domains of the H\textsubscript{II}-phase
  generally form mutually unaligned domains, which 
  %are detrimental to 
  hurts
  membrane permeability. This domain scale misalignment
  had inhibited further development of this technology, and research efforts
  were focused on the Q\textsubscript{I} phase, whose geometry does not require
  alignment~\cite{zhou_new_2007}.

%MRS5: maybe less about Q_I since it's not in this paper?  The introduction should be an introduction to the paper.
%  The most promising Q\textsubscript{I} phase for membrane
%  applications forms at 70\degree C when monomer is mixed with c.a.~20 wt\%
%  glycerol\cite{carter_glycerol-based_2012}. The Q\textsubscript{I} phase
%  consists of a network of three dimensionally interconnected pores (See
%  Figure~\ref{fig:bcc_phases}). This complex geometry effectively adds a degree
%  of tortuosity to the membrane which prevents optimal through-plane transport.
%  Despite the geometric differences between the Q\textsubscript{I} and
%  H\textsubscript{II} phase, the chemical environment within the pores is
%  expected to be similar, and either membrane should be highly selective.

%  % BJC: I like this figure, but I don't know if it is necessary in the main body. i  % Maybe move to supplemental. Thoughts?
%  \begin{figure}
%  \centering
%  \begin{subfigure}{0.45\textwidth}
%        \centering
%        \includegraphics[trim=-2cm 0 0 0, clip, height=4.5cm]{Dibrpyr14.eps}
%        \includegraphics[height=4.5cm]{bcc_phases.png}
%        \caption{}\label{fig:bcc_phases}
%  \end{subfigure}
%  \hspace{1cm}
%  \begin{subfigure}{0.45\textwidth}
%        \centering
%        \includegraphics[trim=0 -0.5cm 0 -0.5cm, clip, height=4.5cm]{NaGA3C11.eps}
%        \includegraphics[height=4.5cm]{hexagonal_packing.png}
%        \caption{}\label{fig:hII_phase}
%  \end{subfigure}
%%MRS5: not sure Q should be in here.  Could cut length by removing some of the detail about Q phase, since it 
%%isn't discussed in the paper. 
%  \caption{The choice of LLC monomer and solvent leads to different accessible liquid
%        crystalline phases. The monomer shown in (a) forms the Q\textsubscript{I} phase
%        in either the Ia3d (left) or Pn3m (right) space group. The surfaces shown
%        represent the interface between the hydrophilic and hydrophobic regions of the unit
%        cell. 
%        %MRS7: not needed
%        %As pictured, we assume that each region occupies an equal volume. 
%        The
%        monomer shown in (b) will form the hexagonal columnar phase
%        (H\textsubscript{II} in the presence of water or Col\textsubscript{h} when no
%         water is present) The blue region represent the hydrophilic pores while the cyan region represents the surrounding hydrophobic monomer tails.}\label{fig:bcc_v_hII}
%  \end{figure}

  %Nanostructured porous membrane materials have become increasingly popular for
  %aqueous separations applications such as desalination and biorefinement because
  %they offer the ability to control pore architecture at the molecular scale,
  %thereby permitting the design of solute-specific separation membranes
  %\cite{humplik_nanostructured_2011}. One can achieve most membrane-based aqueous separations of
  %small molecules using reverse osmosis (RO) or nanofiltration
  %(NF) \cite{van_der_bruggen_review_2003}. While RO and NF have seen many
  %advances in the past few decades, they are far from perfect separation

   %MRS4: these two paragraphs are perhaps not ideal.  They describe the process
  %of making the membranes, but that is not what we need as the introduction to
  %THIS paper.  Instead, you need to focus on the strength/issues with RO and
  %NF that motivate this research.
  % BJC2: The point I am trying to make is that the way the membranes are made 
  % results in either low selectivity or low permeability. I could just state
  % that, or give a little background on the processes to make the point more
  % tangible

  %Commercial RO membranes are typically dense thin-film composite (TFC)
  %membranes with a porous support layer and an active layer made of a polymer
  %matrix formed through interfacial polymerization \cite{jeong_interfacial_2007}.
  %During interfacial polymerization, a step-growth polymerization reaction occurs
  %at the interface of an aqueous monomer solution that has been soaked into the
  %support layer, and an organic solution of a second monomer. The polymer matrix
  %is dense and highly cross-linked which gives high rejections but low
  %permeability.

  %During membrane fabrication,
  %aqueous diamine solution is allowed to penetrate into the polysulfone support
  %layer which is then immersed in trimesoyl chloride, a compound that is
  %immiscible in water. Diamine polymerizes with trimesoyl chloride at the
  %solution interface and creates a thin film that physically adheres to the
  %polysulfone support.

%  solution.  Through the use of a non-solvent, a polymer-rich and polymer-poor
%  phase is formed. A membrane is created as the polymer-poor phase forms pores in
%  the polymer-rich phase. 
%  During phase inversion, a thin polymer film is precipitated out of a polymer

%  NF membranes are typically porous membranes made by a phase-inversion
%  process. The most widely used phase-inversion process is immersion
%  precipitation, during which one submerges a polymer, dissolved in a solvent, in
%  a non-solvent. A solid, porous polymer membrane is all that remains once all
%  solvent has been removed by non-solvent exchange
%  \cite{smolders_microstructures_1992}.  The resultant pores are polydisperse in
%  size, which hurts membrane selectivity. A second technique used to create NF
%  membranes is call track-etching in which a polymer film is bombarded with
%  charged particles, then chemically etched to create pores
%  \cite{apel_track_2001}. The pores are uniform, which benefits selectivity;
%  however, the membranes have a low porosity and subsequently low permeability.

%MRS3: this is a little disconnected to the contents as well.  You want
%the paper to explain the why of doing this research.  Paragraph gives background, but doesn't motivate.

%  Current commercial RO membranes are thin film composite membranes with a
%  porous support layer and an active layer made of a dense polymer matrix. The
%  membranes are unstructured with tortuous and polydisperse diffusion pathways.
%  Separation occurs based on differences in solubility and diffusivity of solutes
%  within the polymer matrix \cite{wijmans_solution-diffusion_1995}. With
%  optimization, one can exploit these differences to create a functional
%  selective barrier. RO operation costs are suboptimal because high feed
%  pressures are required in order to achieve a useful solvent flux.

  %BJC : I don't know that this paragraph fits.  
  %MRS3: not really. Maybe take a look at Osuji's recent review (2016, I think) that addresses applications of nanostructured membranes and uses.  Could rewrite the introduction to be more about the general possibilities of nanostructured membranes, not just desalination.  It should be an introduction to what you will talk about in the paper, not to everything related.
 % Improved RO membranes can make desalination more sustainable. Only 0.5\% of
 % the world's water is fresh. Desalination is an important source of potable
 % water, necessary in order to keep up with demand generated by a rapidly growing
 % global population. Although RO is currently the cheapest form of desalination,
 % the cost must be further decreased in order to meet demand. A typical seawater
 % desalination process requires feed pressures in the range of 800 to 1000 psi
 % \cite{fritzmann_state---art_2007}. By designing better membranes for
 % desalination, we can achieve higher water fluxes and reduce feed pressure
 % requirements which will decrease the cost of fresh water production.
 % \cite{humplik_nanostructured_2011}.

  % financially straining developing regions and contributing strongly to
  % CO\textsubscript{2} emissions. \cite{mcginnis_global_2008} Moreover,
  % designing RO membranes to achieve targeted separations of specific solutes
  % in a complex feed solution is nearly impossible

  %NF was introduced as an intermediate between RO and ultrafiltration, having
  %the ability to separate organic matter and salts on the order of one nanometer
  %in size. Explicit pore pathways running through the thickness of the membrane
  %permit high solvent fluxes at lower pressures than RO. NF membranes typically
  %have a surface charge resulting from ionizable groups which affords separation
  %of ions smaller than the pore radius by the mechanism of Donnan exclusion
  %\cite{van_der_bruggen_review_2003}. NF is often used as a precursor to reverse
  %osmosis since it can efficiently remove a significant portion of dissolved
  %solutes. Following NF pretreatment, RO can further purify the permeate. 
  %Unfortunately, NF membranes, like RO, possess a pore size
  %distribution which limits their ability to perform precise separations
  %\cite{bowen_modelling_2002}.

%  The permeability-selectivity tradeoff has the potential to be overcome by
%  designing membranes at the molecular level. Next-generation nanoporous
%  membranes with high selectivity, permitted by a precisely controlled pore size,
%  and high permeability, allowed by its porous architecture, have the potential to
%  replace traditional RO and NF membrane technologies.

%MRS3: I think that this paragraph probably should be couched in terms of what they COULD do, since are not widely used yet.
%MRS3: if talking about existing membranes, should cite performance.
%  Nanostructured membranes can bypass many of the performance issues which
%  plague traditional NF and RO membranes. One can accomplish targeted separations
%  with high selectivity by tuning shape, size and functionality of the molecular
%  building blocks which form these materials. As a result, solute rejecting pores
%  can have their sizes tuned uniformly, resulting in strict size cut-offs.
%  Entirely different mechanisms may govern transport in a given nanostructured
%  material which can inspire novel separation techniques.

%  Development of nanostructured porous materials has been limited by the
%  ability to synthesize and scale various fundamentally sound technologies.
%  Graphene sheets are atomically thick which results in excellent water
%  permeability but defects during manufacturing severely impact selectivity
%  \cite{cohen-tanugi_multilayer_2016}. Molecular dynamics (MD) simulations of
%  carbon nanotubes show promise \cite{humplik_nanostructured_2011}, but synthetic
%  techniques are unable to achieve scalable alignment and pore monodispersity
%  \cite{hata_water-assisted_2004,maruyama_growth_2005}. Zeolites have sub-nm
%  pores with a narrow pore size distribution, and MD simulations of these
%  materials show that they exhibit complete rejection of solvated ions
%  \cite{murad_molecular_1998}- however, experimental rejection was low and
%  attributed to interstitial defects formed during membrane synthesis
%  \cite{li_desalination_2004}. Consequently, there is a need for a scalable
%  nanostructured porous membrane.

%  Polymeric membranes based on self-assembling lyotropic liquid crystals (LLCs)
%  are suitable candidates for aqueous separation applications. LLCs are
%  amphiphilic molecules that share the characteristic ability of nanostructured
%  porous membrane materials to create highly ordered nanostructures with the
%  added benefits of low cost and synthetic techniques feasible for large-scale
%  production \cite{feng_scalable_2014}. LLC systems created by the monomer
%  Na-GA3C11 (Fig.~\ref{fig:monomer}) have been extensively studied experimentally
%  \cite{feng_scalable_2014,smith_ordered_1997,zhou_supported_2005,resel_h2-phase_2000,feng_thin_2016}.
%  The neat Na-GA3C11 monomer forms the thermotropic liquid crystalline (TLC)
%  Col\textsubscript{h} phase (Fig.\ref{fig:assembly}). The presence of ca. 10
%  wt\% added water results in formation of the lyotropic H\textsubscript{II}
%  phase. In both cases, Na-GA3C11 monomers assembles into mesophases made of
%  hexagonally packed, uniform size, cylinders with hydrophilic head groups
%  oriented inward towards the cylinder center. This LLC assembly is then
%  polymerized in situ to form a cross-linked polymer network to stabilize the
%  structure. The hydrophilic region can act as a pore for aqueous separations
%  \cite{zhou_supported_2005}.  One can envision tailoring the pore region for
%  specific separations by changing the monomer chemistry
%  \cite{resel_h2-phase_2000}.

%  Research into the H\textsubscript{II} phase of polymerized LLC membranes has
%  been revived in recent years. 
%  During early stages of exploration, hexagonal
%  mesophases formed by Na-GA3C11 (Figure~\ref{fig:assembly}) could not be
%  macroscopically aligned, resulting in low flux membranes, and no clear route
%  towards scalable and economical filtration. 
%  For this reason, research efforts

@@

  % BJC: Not sure this figure needs to be in the paper. 
  %MRS2: good question.  It helps give people context, which is necessary to
  %move forward.  Perhaps leave it in for now, though may need to get
  %permission to reprint it.
  %\begin{figure} \centering
  %\includegraphics[width=\linewidth]{soft_confinement.png} \caption{The wedge
  %	  shaped liquid crystal monomer (a) self assembles into mesophases with
%		  hexagonally packed pores (b). The pores are made of stacked monomer disks. A
%		  sub-micron-thick film is created by casting a dilute solution of Na-GA3C11/THF
%		  solution onto a silicon substrate and and allowing the solvent to evaporate.
%		  The thin film contains nanoporous columns which lie parallel to the film plane.
%		  (d) When a soft PDMS pad is imposed to the thin film, with subsequent thermal
%		  annealing, the columns align perpendicular to the film plane. (e)
%		  Photo-cross-linking of the aligned film creates a mechanically stable thin film
%		  with vertically aligned nanopores}~\label{fig:soft} \end{figure}

  \begin{figure}
	\centering
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{NaGA3C11.pdf}
		\caption{}~\label{fig:monomer}
	\end{subfigure}
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{monomer_twocolor.pdf}
		\caption{}~\label{fig:atomistic_monomer}
	\end{subfigure}
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[width=\textwidth]{wedge_thick.pdf}
		\caption{}~\label{fig:wedge}
	\end{subfigure}
		\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{columns.pdf}
		\caption{}~\label{fig:wedge_layer}
	\end{subfigure}
	\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{hexagonal_packing.pdf}
		\caption{}~\label{fig:hex_packing_simple}
	\end{subfigure}
        %MRSF: in the actual paper, there would only be room for about a half-page figure, max.  Can hyou reduce the especially the c/d/e sizes so that it fully fits in a half-page figure? 
	\caption{(a) The LLC monomer Na-GA3C11 (b) rendered atomistically (c)
	exhibits wedge-like character. (d) Monomers stack on top of each other to
	create columns with short range order then assemble into pores with  
	hydrophilic head groups (blue) facing towards the pore center. (e) The 
	pores assemble into hexagonally packed columnar mesophases.}~\label{fig:assembly}
  \end{figure}

  % BJC14: (5) justify need for simulations

@@

  A molecular-level understanding of LLC polymer membrane structure, enabled by
  molecular dynamics (MD) simulations, can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. Useful molecular-level modeling should incorporate a detailed picture of the
  nanoscopic pore structure, which is crucial to understanding the role of
  monomer structure in solute transport and membrane design. 
  Atomistic molecular dynamics simulations can provide the required level of detail
  (Figure \ref{fig:detail}), assuming the force fields are sufficiently accurate.
  With such an atomistic model, we can directly observe molecular-level solute
  transport and suggest governing mechanisms. We can also observe how the choice of
  head group interacts with solutes of interest. In addition, we can interchange
  counterions which may influence both the pore size and the strength of the
  Donnan potential. 
  %which affects the degree to which the membrane can exclude
  %charged species at the membrane surface.

@@

  %MRS8: I started tweaking the below, but it might be better just to eliminate.
  % Modeling of Col\textsubscript{h} will help in modeling 
  %H\textsubscript{II} in the future.
  %The two phases appear to share similar structural characteristics since the
  %pore spacings in each system are in close agreement
  %\cite{feng_thin_2016,resel_structural_2000}.

%  We used experimental small-angle X-ray scattering (SAXS) data from
%  \cite{feng_thin_2016} (Fig.~\ref{fig:SAXS}) and wide angle X-ray scattering
%  (WAXS) data (Fig.~\ref{fig:WAXS}, produced as described in
%  \cite{feng_scalable_2014}) for comparison to our model. We rely primarily on the 2D WAXS data
%  since it encodes all structural details down to the sub-nm scale.  
  There are five major features of interest present in the 2D experimental
  pattern shown in Figure~\ref{fig:WAXS}.

@@

  \newpage
  %BJC: TOC graphic

>>>>

\title{Understanding the Nanoscale Structure of Inverted Hexagonal Phase Lyotropic Liquid Crystal Polymer Membranes}
\author{Benjamin J. Coscia}
\affiliation{Department of Chemical and Biological Engineering, University of Colorado Boulder, Boulder, CO 80309, USA}
\author{Joseph Yelk}
\author{Matthew A. Glaser}
\affiliation{Department of Physics, University of Colorado Boulder, Boulder CO, 80309, USA}
\author{Douglas L. Gin}
\affiliation{Department of Chemical and Biological Engineering, University of Colorado Boulder, Boulder, CO 80309, USA}
\alsoaffiliation{Department of Chemistry and Biochemistry, University of Colorado Boulder, Boulder CO 80309, USA}
\author{Xunda Feng}
\affiliation{Department of Chemical and Environmental Engineering, Yale University, New Haven, Connecticut 06511, USA}
\author{Michael R. Shirts}
\email{michael.shirts@colorado.edu}
\affiliation{Department of Chemical and Biological Engineering, University of Colorado Boulder, Boulder, CO 80309, USA}
\begin{document}

  \graphicspath{{./figures/}}

@@

  %NOTE: all deleted comments are still in Draft.tex

@@

  \end{abstract}

@@

  More highly selective nanoporous membranes would be extremely useful 
  in performing complex aqueous separations with seawater and various 
  types of wastewater. For example, Sodium chloride and boron in seawater
  \cite{fritzmann_state---art_2007} and organic micropollutants found in
  municipal and industrial wastewaters \cite{schwarzenbach_challenge_2006} represent
  just a few of the diverse contaminants of water sources. By efficiently
  separating contaminants from feed solutions with highly selective membranes,
  it is possible to reduce the number of required membrane passes and post-treatment steps
  needed for a given filtration process \cite{werber_materials_2016}, thus
  lowering energy requirements. Additionally, one can also extract valuable resources
  from the feed streams. For example, flowback water produced during hydraulic
  fracturing of shale formations contains dissolved species such as acetate whose
  extraction has economic value \cite{dischinger_application_2017}.

  Reverse osmosis (RO) and nanofiltration (NF) are two prevailing membrane
  filtration processes that can be used to separate solutes on the order of 1 nm
  in size and smaller, including 
  %MRSF: this does seem to blend RO and NF together; is this correct, or are there important differences that should be outlined?
  %BJCF: tried to differentiate them
  ions. However, commercially available amorphous RO membranes and porous NF
  membranes lack uniformity, which severely impacts their
  selectivity \cite{van_der_bruggen_review_2003}. Although scalable, their
  fabrication involves the spontaneous assembly of polymers into disordered
  structures that offer separation pathways that are tortuous and polydisperse
  in size \cite{werber_materials_2016}. Tortuosity increases the effective
  length that a solute must travel, while pore size polydispersity limits membrane
  selectivity.

@@

@@

  The H\textsubscript{II}-phase pore geometry (Figure~\ref{fig:assembly}) has a
  higher theoretical capacity for transport than the Q\textsubscript{I} phase.
  The H\textsubscript{II} phase forms at room temperature in the presence of
  ca.~10 wt\% water and consists of hexagonally packed, hydrophilic pore
  columns\cite{smith_ordered_1997}. In the absence of water, neat monomer will
  form the same hexagonal columnar structure which, in the literature, has been
  referred to as the Col\textsubscript{h} thermotropic phase\cite{feng_scalable_2014}.
  Q\textsubscript{I}-phase membranes consist of a tortuous network of three
  dimensionally interconnected pores that prevent optimal through-plane
  transport. In contrast, the densely packed, non-tortuous and uniform sized
  pores of H\textsubscript{II}-phase membranes represent the ideal geometry for
  achieving high solute flux\cite{matyka_tortuosity-porosity_2008}.  However, the
  hexagonally packed LC domains of the H\textsubscript{II}-phase
  generally form mutually unaligned domains, which hurts membrane permeability.
  This domain scale misalignment had inhibited further development of this technology,
  and research efforts were focused on the Q\textsubscript{I} phase, whose geometry
  does not require alignment~\cite{zhou_new_2007}.

@@

  \begin{figure}
	\centering
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{NaGA3C11.pdf}
		\caption{}~\label{fig:monomer}
	\end{subfigure}
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{monomer_twocolor.pdf}
		\caption{}~\label{fig:atomistic_monomer}
	\end{subfigure}
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[width=0.6\textwidth]{wedge_thick.pdf}
		\caption{}~\label{fig:wedge}
	\end{subfigure}
		\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=0.6\textwidth]{columns.pdf}
		\caption{}~\label{fig:wedge_layer}
	\end{subfigure}
	\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=0.6\textwidth]{hexagonal_packing.pdf}
		\caption{}~\label{fig:hex_packing_simple}
	\end{subfigure}
	\caption{(a) The LLC monomer Na-GA3C11 (b) rendered atomistically (c)
	exhibits wedge-like character. (d) Monomers stack on top of each other to
	create columns with short range order then assemble into pores with  
	hydrophilic head groups (blue) facing towards the pore center. (e) The 
	pores assemble into hexagonally packed columnar mesophases.}~\label{fig:assembly}
  \end{figure}

@@

  A molecular-level understanding of LLC polymer membrane structure, enabled by
  molecular dynamics (MD) simulations, can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. Useful molecular-level modeling should incorporate a detailed picture of the
  nanoscopic pore structure, which is crucial to understanding the role of
  monomer structure in solute transport and membrane design. 
  Atomistic molecular dynamics simulations can provide the required level of detail
  (Figure \ref{fig:detail}), assuming the force fields are sufficiently accurate.
  With such an atomistic model, we can directly observe molecular-level solute
  transport and suggest governing mechanisms. We can also observe how the choice of
  head group interacts with solutes of interest. In addition, we can interchange
  counterions which may influence both the pore size and the strength of the
  Donnan potential.

@@

  There are five major features of interest present in the 2D experimental
  pattern shown in Figure~\ref{fig:WAXS}.

@@

  \newpage
